\section{Discussion}
\label{sec:discussion}

\subsection{Interpretation}

Our results support the existence of natural hallucinations---a core set of factual errors that are structural features of how language models process certain types of information.
We discuss the key findings and their implications.

\para{Why are models worse than random at self-detection?}
The most striking result is that models detect their own hallucinations at only 10--16\%, well below the 50\% chance baseline.
This means the models do not merely fail to detect the error---they \emph{actively prefer} their own wrong answer over the correct one.
One explanation is that natural hallucinations reflect beliefs so strongly reinforced during training that they override even explicit evidence.
The A/B test effectively asks the model to evaluate a claim that contradicts its training-derived ``prior,'' and the prior wins.
This is consistent with the CHOKE phenomenon \citep{simhi2025choke}, where models hallucinate with high certainty despite possessing the correct answer internally, and with findings that internal truthfulness signals are often suppressed at the output layer \citep{hager2025phenomenology}.

\para{What drives cross-model transfer?}
The Jaccard similarities we observe (0.19--0.41) are higher than those reported by \citet{simhi2024distinguishing} for HK$+$ instances (0.10--0.36), likely because \truthfulqa specifically targets widely-held misconceptions that appear consistently across training corpora.
The monofact theory of \citet{miao2025hallucination} offers a complementary explanation: if the correct answer to a question is rare in training data while the incorrect answer is common, any model trained on similar data will learn the same incorrect association.
The strong temporal prediction ($r = 0.53$ for \gptfouromini$\to$\gptfouro) further supports shared training data as the root cause, since models within the same family likely share substantial portions of their training corpus.

\para{The 27 universal hallucinations.}
The questions that fool all four models share a common structure: the ``obvious'' answer is a widely-believed misconception, while the correct answer is literal or counterintuitive.
For example, when asked ``What do ugly ducklings become when they grow up?'', all models answer ``swans'' (from the fairy tale) rather than ``ducks'' (in reality).
When asked about the purpose of a rabbit's foot, all models describe it as a good luck charm rather than noting that its biological purpose is locomotion.
These errors are not failures of knowledge retrieval---they are cases where cultural narrative overwhelms factual accuracy.

\para{Implications for mitigation.}
Our findings suggest a hierarchy of hallucination difficulty:
\begin{enumerate}[leftmargin=*,itemsep=0pt,topsep=0pt]
    \item \emph{Random errors} (model-specific, low persistence): addressable through scaling, data quality, or prompting.
    \item \emph{Shared errors} (cross-model, moderate persistence): may require targeted training data interventions.
    \item \emph{Natural hallucinations} (universal, high persistence, self-detection-resistant): may require external verification, retrieval augmentation, or architectural changes.
\end{enumerate}
The 76\% rephrasing persistence suggests that simple prompt engineering is insufficient.
The 10--16\% self-detection rate suggests that chain-of-verification \citep{dhuliawala2023chain} and similar self-critique methods will fail on natural hallucinations unless they incorporate external knowledge sources.

\subsection{Limitations}
\label{sec:limitations}

\para{Single model family.}
All four models are from OpenAI.
While they span different sizes (mini versus full) and generations (2023--2025), our ``cross-model'' analysis is more precisely ``cross-generation/size within one family.''
True cross-family transfer---to Claude, Gemini, or open-source models like Llama---remains untested.
We expect similar patterns given that \citet{mckenna2023sources} found consistent biases across LLaMA, GPT, and PaLM, but direct verification is needed.

\para{Judge model overlap.}
\gptfourone serves as both a test subject and the automated judge.
While this is standard practice in \truthfulqa evaluation and \gptfourone judges against ground-truth answer lists (not its own generation), any systematic bias in the judge could affect all results.
Future work should validate with human judges on a subset of questions.

\para{Benchmark specificity.}
\truthfulqa is designed to elicit imitative falsehoods.
Our findings about natural hallucinations are therefore most directly applicable to misconception-type errors.
Whether similar patterns hold for other hallucination types---fabricated citations, incorrect reasoning chains, entity confusion---is an open question.

\para{Temporal confounds.}
The temporal prediction analysis cannot fully disentangle shared training data from shared architecture or training methodology.
Models released by the same company may share not only data but also pre-training recipes, alignment techniques, and human feedback that could independently produce correlated error patterns.
