\section{Introduction}
\label{sec:intro}

Large language models produce factual errors.
Despite rapid progress in scale, training data, and alignment techniques, hallucinations remain a persistent failure mode across model families and generations \citep{lin2022truthfulqa, huang2023survey}.
A common assumption is that these errors are essentially random---artifacts of insufficient training data or imperfect optimization that will gradually disappear as models improve.
But what if some hallucinations are not random at all?

We investigate this question by studying what we call \emph{natural hallucinations}: factual errors that multiple models consistently produce, that survive question rephrasing, and that models cannot recognize as wrong even when presented with the correct answer side by side.
If such hallucinations exist, they represent a qualitatively different challenge from random errors---one that scaling alone is unlikely to solve.

\para{Why this matters.}
The distinction between random and natural hallucinations has direct practical consequences.
Random errors can be addressed through better data curation, larger models, or straightforward prompting strategies.
Natural hallucinations, by contrast, may require fundamentally different interventions: targeted training data corrections, architectural changes, or external verification systems.
Understanding which hallucinations are natural is therefore critical for both safety research and deployment decisions.

\para{Gap in existing work.}
Prior work has established that LLMs produce ``imitative falsehoods'' that match common human misconceptions \citep{lin2022truthfulqa}, that some hallucinations persist despite the model encoding correct information internally \citep{orgad2024llmsknow, simhi2025choke}, and that training data frequency drives hallucination rates \citep{miao2025hallucination}.
However, no study has systematically characterized the \emph{same set of questions} across all four dimensions: cross-model transfer, robustness to rephrasing, failure of self-detection, and temporal prediction from older to newer models.

\para{Our approach.}
We test four OpenAI models---\gptfourone, \gptfouro, \gptfouromini, and \gptthreefive---on all 817 \truthfulqa questions.
We identify questions that cause failures across multiple models, then probe these shared errors through three follow-up experiments: rephrasing robustness, self-detection with evidence, and cross-model transfer analysis.
Our results paint a consistent picture: a core set of hallucinations behaves as a structural feature of language modeling, not as a bug to be patched.

\para{Key findings.}
We find that 27 questions (3.3\%) cause all four models to hallucinate, with 82 questions (10.0\%) fooling three or more.
These hallucinations persist at a 76\% rate across five rephrasings per question.
In an A/B self-detection test, models choose the correct answer only 10--16\% of the time---worse than random---while achieving 66--95\% on control questions where they originally answered correctly.
Cross-model Jaccard similarity ranges from 0.19 to 0.41 (all $p < 0.001$), and \gptfouromini hallucinations predict \gptfouro hallucinations with $r = 0.53$.

In summary, our main contributions are:
\begin{itemize}[leftmargin=*,itemsep=0pt,topsep=0pt]
    \item We identify 27 universal and 82 near-universal hallucinations that persist across four models spanning three years of development, establishing the existence of natural hallucinations.
    \item We demonstrate that these errors are robust to rephrasing (76\% persistence) and resistant to self-detection (10--16\% accuracy, $z = 18.49$, $p < 0.0001$), revealing a systematic bias rather than a surface-level pattern match.
    \item We show that hallucination sets overlap 3--6$\times$ above random across model pairs and that older-model errors significantly predict newer-model errors ($r = 0.26$--$0.53$), suggesting shared training data biases as a root cause.
\end{itemize}
